The data.table Command
======================

```{r}
library(data.table)
```

A data.frame is created like this:
```{r}
set.seed(0)
df = data.frame(x = rnorm(9), y = rep(letters[1:3], each = 3), z = rnorm(9))
print(df)
```

A data.table is created in an identical fashion:
```{r}
set.seed(0)
dt <- data.table(x = rnorm(9), y = rep(letters[1:3], each = 3), z = rnorm(9))
print(df)
```

### Subsetting

Here are some subsetting examples (same as with data.tables):
```{r}
dt[2, ]
dt[dt$y == "a", ]
dt[c(2, 3), ]
```

### Expression

Trying to subset on columns in data tables is **not** the same as with data frames!  We actually specifying expressions rather than subset indices.  For example, if you specify a list, that list of functions will be performed:
```{r}
    dt[, list(mean(x), sum(z))]
```

### New Columns
This can be done using a colon:
```{r}
    dt[, w:=z^2]
    dt[, m := {tmp <- (x + z); log2(tmp + 5)}]
    dt[, a := x > 0]
```

### Speedy Operations
This is very quick
```{r}
    set.seed(123)
    dt <- data.table(x = sample(letters[1:3], 1e5, TRUE))
    dt[, .N, by = x]
```

### Setting Keys
Keys can be set like this
```{r}
    dt <- data.table(x = rep(letters[1:3], each = 100), y = rnorm(300))
    dt[dt$x == 'a']
    setkey(dt, x)
    dt['a']
```

### Merging Data
Here is an example of data merging.  Merging forms a merged table based on common keys.  This is faster than data.tables
```{r}
    DT1 <- data.table(x = c('a', 'a', 'b', 'dt1'), y=1:4)
    DT2 <- data.table(x = c('a', 'b', 'b', 'dt2'), z = 5:8)
    setkey(DT1, x); setkey(DT2, x)
    merge(DT1, DT2)
```

### Reading/writing
Also, it is much faster to read and write to data.tables then data.frames.  it is well worth looking into this.
